---
title: "Simple cluster"
author: "J. Meijer"
date: "2025-01-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```




Check for weird numbers:
- chrfigs_fam_age
- chrchs_timeslept



Check if variables with only one option are removed.


```{r}
# Load libraries
library(flextable)
library(officer)
library(dplyr)
library(readr)
library(readxl)

```

# Create and load basetable

```{r, tidy=TRUE}
#Load function
source("table_helpers/Basetable_function.R")

# Set path to dictionary
dictionary_DIR <- "path/to/data_dictionary.xlsx"

#Set path to data
data_DIR <- file.path( "path/to/restricted_data")

#Run basetable function. 
results <- create_basetable(
  data_DIR = data_DIR,
  dictionary_DIR = dictionary_DIR,
  release = "Amp_scz4",
  timepoint= "baseline"
)

data <- results$data
meta <- results$meta

 
results_screening <- create_basetable(
  data_DIR = data_DIR,
  dictionary_DIR = dictionary_DIR,
  release = "Amp_scz4",
  timepoint= "screening"
)

data_screening <- results_screening$data
meta_screening <- results_screening$meta



results_month1 <- create_basetable(
  data_DIR = data_DIR,
  dictionary_DIR = dictionary_DIR,
  release = "Amp_scz4",
  timepoint= "month_1"
)
data_month1 <- results_month1$data
meta_moth1 <- results_month1$meta

results_month2 <- create_basetable(
  data_DIR = data_DIR,
  dictionary_DIR = dictionary_DIR,
  release = "Amp_scz4",
  timepoint= "month_2"
)
data_month2 <- results_month2$data
meta_moth2 <- results_month2$meta


results_month3<- create_basetable(
  data_DIR = data_DIR,
  dictionary_DIR = dictionary_DIR,
  release = "Amp_scz4",
  timepoint= "month_3"
)
data_month3 <- results_month3$data
meta_moth3 <- results_month3$meta

results_month4 <- create_basetable(
  data_DIR = data_DIR,
  dictionary_DIR = dictionary_DIR,
  release = "Amp_scz4",
  timepoint= "month_4"
)
data_month4 <- results_month4$data
meta_moth4 <- results_month4$meta

results_month5 <- create_basetable(
  data_DIR = data_DIR,
  dictionary_DIR = dictionary_DIR,
  release = "Amp_scz4",
  timepoint= "month_5"
)
data_month5 <- results_month5$data
meta_moth5 <- results_month5$meta



```

```{r}

## 1. Add any columns that exist only in screening ----------------------------
new_cols   <- setdiff(names(data_screening), names(data))
merged_data <- data %>%                     # baseline rows are the backbone
  left_join(select(data_screening, src_subject_id, all_of(new_cols)),
            by = "src_subject_id")

## 2. Decide, variable-by-variable, whether to keep baseline or screening -----
common_vars <- setdiff(intersect(names(data), names(data_screening)),
                       "src_subject_id")    # overlap apart from the key

# Count how many non-missing / non-empty values each data set has
baseline_n  <- sapply(common_vars,
                      function(v) sum(!is.na(data[[v]]) & data[[v]] != ""))
screening_n <- sapply(common_vars,
                      function(v) sum(!is.na(data_screening[[v]]) &
                                      data_screening[[v]] != ""))

vars_take_screening <- names(screening_n)[screening_n > baseline_n]

## 3. Overwrite those columns *row-wise* with screening values ---------------
if (length(vars_take_screening) > 0) {
  merged_data[vars_take_screening] <-
    data_screening %>%
      select(src_subject_id, all_of(vars_take_screening)) %>%
      right_join(select(merged_data, src_subject_id), by = "src_subject_id") %>%
      select(all_of(vars_take_screening))
}

# merged_data now contains:
#   • baseline values everywhere baseline is as complete or more complete
#   • screening values for columns where screening had strictly fewer missing values
```

### Remove useless columns
```{r}
## Remove columns with only 1 unique value
merged_data <- merged_data[, sapply(merged_data, function(col) length(unique(col))>1)]

## Data preprocessing: Traumatic brain injury

# Combine age with yes or no column to get rid of the Non applicable (-3)
if("chrtbi_age_inj" %in% colnames(merged_data)){
  merged_data$chrtbi_age_inj <- paste(merged_data$chrtbi_subject_head_injury, merged_data$chrtbi_subject_age)
  merged_data$chrtbi_age_inj[merged_data$chrtbi_age_inj=="No NA"] <- "No"
  merged_data$chrtbi_age_inj[merged_data$chrtbi_age_inj=="NA NA"] <- NA
}


# Combine how many times with yes or no column to get rid of the Non applicable (-3)
if("chrtbi_ntimes" %in% colnames(merged_data)){
  merged_data$chrtbi_ntimes <- paste(merged_data$chrtbi_subject_head_injury, merged_data$chrtbi_subject_times)
  merged_data$chrtbi_ntimes[merged_data$chrtbi_ntimes=="No NA"] <- "No"
  merged_data$chrtbi_ntimes[merged_data$chrtbi_ntimes=="NA NA"] <- NA
}
# Combine how many times with yes or no column to get rid of the Non applicable (-3)
if("chrtbi_symptoms" %in% colnames(merged_data)){
  merged_data$chrtbi_symptoms <- paste(merged_data$chrtbi_subject_head_injury, merged_data$chrtbi_subject_symptoms)
  merged_data$chrtbi_symptoms[merged_data$chrtbi_symptoms=="No NA"] <- "No"
  merged_data$chrtbi_symptoms[merged_data$chrtbi_symptoms=="NA NA"] <- NA
}
# Combine how many times with yes or no column to get rid of the Non applicable (-3)
if("chrtbi_knockedout" %in% colnames(merged_data)){
  merged_data$chrtbi_knockedout <- paste(merged_data$chrtbi_subject_head_injury, merged_data$chrtbi_subject_knockedout)
  merged_data$chrtbi_knockedout[merged_data$chrtbi_knockedout=="No NA"] <- "No"
  merged_data$chrtbi_knockedout[merged_data$chrtbi_knockedout=="NA NA"] <- NA
}

#Remove unneeded columns
if("chrtbi_age_inj" %in% colnames(merged_data)){
  merged_data <- select(merged_data, -c(chrtbi_subject_head_injury, chrtbi_subject_age, chrtbi_subject_times,
                                        chrtbi_subject_symptoms, chrtbi_subject_knockedout))
}

## Include these new variables in the meta data

meta <- rbind(meta, data.frame(
  ElementName = 'chrtbi_age_inj',
  Datatype = 'Traumatic Brain Injury',
  Modality = 'Brain_injury',
  stringsAsFactors = FALSE
))

meta <- rbind(meta, data.frame(
  ElementName = 'chrtbi_ntimes',
  Datatype = 'Traumatic Brain Injury',
  Modality = 'Brain_injury',
  stringsAsFactors = FALSE
))

meta <- rbind(meta, data.frame(
  ElementName = 'chrtbi_symptoms',
  Datatype = 'Traumatic Brain Injury',
  Modality = 'Brain_injury',
  stringsAsFactors = FALSE
))

meta <- rbind(meta, data.frame(
  ElementName = 'chrtbi_knockedout',
  Datatype = 'Traumatic Brain Injury',
  Modality = 'Brain_injury',
  stringsAsFactors = FALSE
))

# If variable is about a medication dose, change all -3 and -300 to 0
# Loop through columns that have 'dose' in their names (case-insensitive)
for (col in names(merged_data)) {
  if (grepl("dose", col, ignore.case = TRUE)) {
    merged_data[[col]][merged_data[[col]] %in% c(-3, -300)] <- 0
  }
}
# Loop through columns that have 'med' in their names (case-insensitive)
for (col in names(merged_data)) {
  if (grepl("med", col, ignore.case = TRUE)) {
    merged_data[[col]][merged_data[[col]] %in% c(-3, -300)] <- 0
  }
}


# Remove all -3 and -300 when possible
merged_data[merged_data == -3 | merged_data == -300] <- NA
data_month2[data_month2 == -3 | data_month2 == -300] <- NA

# names(merged_data)[sapply(merged_data, function(x) any(x %in% c(-3,-300), na.rm=TRUE))] #To test which columns are still containing -3 and -300.


```



### Cognition

#### Standardizing using HC
```{r}

penncnb_vars <- grep("^cnb", names(merged_data), value = TRUE)

# Filter HC group
hc_data <- merged_data[merged_data$phenotype == "HC", ]

# Function to compute z-scores based on HC stats
z_score_hc <- function(x, varname) {
  m <- mean(hc_data[[varname]], na.rm = TRUE)
  s <- sd(hc_data[[varname]], na.rm = TRUE)
  return((x - m) / s)
}

# Apply z-scoring to all identified variables
for (var in penncnb_vars) {
  merged_data[[var]] <- z_score_hc(merged_data[[var]], var)
}
```



## Combine CAARMS 1 and 2 into one score
```{r}
merged_data$CAARMS_delusions <- (merged_data$psychs_caarms_p1+merged_data$psychs_caarms_p2)/2

meta_newrow <- data.frame(ElementName="CAARMS_delusions", Datatype="CAARMS aggregated delusions", Modality="Include_cluster")
meta <- rbind(meta, meta_newrow)

#Remove the individual items from the data
merged_data <- merged_data[, !names(merged_data) %in% c("psychs_caarms_p1", "psychs_caarms_p2")]

```

## Create trauma subscores

Recoding
```{r}
# Normal-coded CTQ items
ctq_normal_vars <- c(
  "chrpps_doc", "chrpps_bruise", "chrpps_belt", "chrpps_physab", "chrpps_beat",
  "chrpps_touch", "chrpps_threat", "chrpps_sexual", "chrpps_molest", "chrpps_sexab",
  "chrpps_lazy", "chrpps_born", "chrpps_hate", "chrpps_hurt", "chrpps_emoab",
  "chrpps_hunger", "chrpps_pardrunk", "chrpps_dirty"
)

# Already reverse-coded CTQ items
ctq_reverse_vars <- c(
  "chrpps_loved",
  "chrpps_special",
  "chrpps_care",
  "chrpps_closefam",
  "chrpps_support",
  "chrpps_protect",
  "chrpps_docr"
)

recode_ctq_normal <- function(x) {
  x <- as.character(x)
  x <- trimws(x)

  dplyr::case_when(
    x %in% c("1", "Never true") ~ 1,
    x %in% c("2", "Rarely true") ~ 2,
    x %in% c("3", "Sometimes true") ~ 3,
    x %in% c("4", "Often true") ~ 4,
    x %in% c("5", "Very often true/Always", "Very often true", "Always") ~ 5,
    x %in% c("-99", "N/A", "NA", "Refused", "77", "Missing", "88") ~ NA_real_,
    TRUE ~ NA_real_
  )
}

recode_ctq_reverse <- function(x) {
  x <- as.character(x)
  x <- trimws(x)

  dplyr::case_when(
    x %in% c("5", "Never true") ~ 5,
    x %in% c("4", "Rarely true") ~ 4,
    x %in% c("3", "Sometimes true") ~ 3,
    x %in% c("2", "Often true") ~ 2,
    x %in% c("1", "Very often true/Always", "Very often true", "Always") ~ 1,
    x %in% c("-99", "N/A", "NA", "Refused", "77", "Missing", "88") ~ NA_real_,
    TRUE ~ NA_real_
  )
}

# Apply correct recoding to each group
merged_data[ctq_normal_vars] <- lapply(
  merged_data[ctq_normal_vars],
  recode_ctq_normal
)

merged_data[ctq_reverse_vars] <- lapply(
  merged_data[ctq_reverse_vars],
  recode_ctq_reverse
)
```


```{r}

merged_data$ctq_physical <- rowSums(
  merged_data[c(
    "chrpps_doc",
    "chrpps_bruise",
    "chrpps_belt",
    "chrpps_physab",
    "chrpps_beat"
  )],
  na.rm = FALSE
)

merged_data$ctq_sexual <- rowSums(
  merged_data[c(
    "chrpps_touch",
    "chrpps_threat",
    "chrpps_sexual",
    "chrpps_molest",
    "chrpps_sexab"
  )],
  na.rm = FALSE
)

merged_data$ctq_emotional <- rowSums(
  merged_data[c(
    "chrpps_lazy",
    "chrpps_born",
    "chrpps_hate",
    "chrpps_hurt",
    "chrpps_emoab"
  )],
  na.rm = FALSE
)

merged_data$ctq_other <- rowSums(
  merged_data[c(
    "chrpps_loved",
    "chrpps_special",
    "chrpps_care",
    "chrpps_closefam",
    "chrpps_support",
    "chrpps_hunger",
    "chrpps_protect",
    "chrpps_pardrunk",
    "chrpps_dirty",
    "chrpps_docr"
  )],
  na.rm = FALSE
)


```

Add to meta table
```{r}
# Add new CTQ aggregated scores to meta
meta_newrows <- data.frame(
  ElementName = c(
    "ctq_physical",
    "ctq_sexual",
    "ctq_emotional",
    "ctq_other"
  ),
  Datatype = c(
    "CTQ aggregated physical abuse",
    "CTQ aggregated sexual abuse",
    "CTQ aggregated emotional abuse",
    "CTQ aggregated other childhood adversity"
  ),
  Modality = c(
    "compare_clusters",
    "compare_clusters",
    "compare_clusters",
    "compare_clusters"
  )
)

meta <- rbind(meta, meta_newrows)

# Remove individual CTQ items from the data
ctq_item_vars <- c(
  "chrpps_doc", "chrpps_bruise", "chrpps_belt", "chrpps_physab", "chrpps_beat",
  "chrpps_touch", "chrpps_threat", "chrpps_sexual", "chrpps_molest", "chrpps_sexab",
  "chrpps_lazy", "chrpps_born", "chrpps_hate", "chrpps_hurt", "chrpps_emoab",
  "chrpps_loved", "chrpps_special", "chrpps_care", "chrpps_closefam",
  "chrpps_support", "chrpps_hunger", "chrpps_protect", "chrpps_pardrunk",
  "chrpps_dirty", "chrpps_docr"
)

merged_data <- merged_data[, !names(merged_data) %in% ctq_item_vars]
```



### Save preprocessed data

```{r}
#Save complete merged_data. NOTE: Change to merged_data_CT if you want to include MRI
write_csv(merged_data, 'path/to/simpleclust_data/merged_data_simpleclust.csv')
write_csv(meta, 'path/to/simpleclust_data/merged_meta_simpleclust.csv')
write_csv(data_month1, 'path/to/simpleclust_data/data_month1_simpleclust.csv')
write_csv(data_month2, 'path/to/simpleclust_data/data_month2_simpleclust.csv')
write_csv(data_month3, 'path/to/simpleclust_data/data_month3_simpleclust.csv')
write_csv(data_month4, 'path/to/simpleclust_data/data_month4_simpleclust.csv')
write_csv(data_month5, 'path/to/simpleclust_data/data_month5_simpleclust.csv')



```

# Demographic table full sample

```{r}
# Source the demographics function
source("table_helpers/Demographictable_function.R")

comparison <- 'phenotype'

comparison_labels <- c("CHR" = "CHR", "HC" = "CC")

# Generate summary table
table <- demographic_table(merged_data, dictionary_DIR, comparison, comparison_labels)

# Print table
print(table)

# Save in a word document
save_as_docx(table, path = "Dem_table_fullSample_simpleclust.docx")
```

# Demographic table final sample with comparison between discovery and test

Load in prescient IDs and add a column sample 
```{r}
# 1. Load the Prescient ID list once:

prescient <- read.delim(
  "path/to/restricted_data/Prescient_Client List_ DPACC ID with RA Name (ALL sites)_20_03_2025.txt",
  sep = "\t",
  header = TRUE,
  stringsAsFactors = FALSE
)

colnames(prescient) <- c(
  "fkLocationID",
  "pkclientid",
  "src_subject_id",
  "RAname"
)

prescient_ids <- unique(prescient$src_subject_id)

disc_data <- merged_data

disc_data$sample <- ifelse(disc_data$src_subject_id %in% prescient_ids, yes="Discovery", no="Validation")

```



```{r}
comparison <- 'sample'

table <- demographic_table(disc_data, dictionary_DIR, comparison)

print(table)
save_as_docx(table, path = "path/to/simpleclust_data/Dem_table_sample_simpleclust.docx")

```

